Quantifying Uncertainty in Forecast-Based Anticipatory Action Triggers Using Bootstrapping Methods

### 2025-09-04 paper.tex
latest text update

To test how sensitive the protocol is to the exact threshold values chosen, this section presents an uncertainty analysis. This analysis is based on the method of re-estimating the trigger threshold with resampled draws of the historical record. Repeating this procedure many times gives us an estimated distribution of how much the threshold might be expected to vary due to sampling error. In this case, we run 1,000 iterations of the re-sampling process. In each iteration, we draw a sample of equal size to the original data comprised of randomly sampled blocks of 5 consecutive years, sampled with replacement. The block width of 5 was chosen here to account for the timescale of common sources of interannual correlation like ENSO, but could be adjusted by the user as they wish.

@@

If the current forecast is outside of the sensitivity range, this indicates that we have statistically robust evidence that it is above or below the threshold. If the current forecast falls within the sensitivity range, this indicates that we should interpret the trigger result for the current season with caution.

>>>>

To test how sensitive the protocol is to the exact threshold values chosen, this section presents an uncertainty analysis. This analysis is based on the method of re-estimating the trigger threshold with resampled draws of the historical record, known as bootstrappaing [cite]. Repeating this procedure many times gives us an estimated distribution of how much the threshold might be expected to vary due to sampling error of a limited historical record. In this case, we run 1,000 iterations of the re-sampling process. In each iteration, we draw a sample of equal size to the original data comprised of randomly sampled blocks of 5 consecutive years, sampled with replacement. The block width of 5 was chosen here to account for the timescale of common sources of interannual correlation like ENSO.

@@

If the current forecast is outside of the sensitivity range, this indicates that we have statistically robust evidence that the forecasted conditions are above or below the threshold. If the current forecast falls within the sensitivity range, this indicates that we should interpret the trigger result for the current season with caution as the appearance of trigger-worth conditions could be occurring just by chance.
